\documentclass[11pt, twoside]{report}

\usepackage{fontspec}
\usepackage[utf8]{inputenc}
\usepackage[bitstream-charter]{mathdesign}
\usepackage{bbding}
\usepackage{ragged2e}
\usepackage{parskip}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{paracol}
\usepackage{mdframed}
\usepackage[margin=1in]{geometry}
\usepackage[autocompile]{gregoriotex}

\titleformat{\chapter}[block]{\huge\scshape\filcenter}{}{1em}{}
\titleformat{\section}[block]{\Large\bfseries}{}{1em}{}

\newcommand{\rubric}[1]{
	\switchcolumn[0] {
		\itshape
		#1
	}
}

\newcommand{\latinenglish}[2]{
	\switchcolumn[0]* {
		#1
	}
	\switchcolumn[1] {
		\itshape\small
		#2
	}
}

\newcommand{\latinenglishequal}[2]{
	\switchcolumn[0]* {
		#1
	}
	\switchcolumn[1] {
		\itshape
		#2
	}
}

\newenvironment{latinenglishsection}
	{\columnratio{.7, .3} \begin{paracol}{2}}
	{\end{paracol}}

\newenvironment{latinenglishequalsection}
	{\columnratio{.5, .5}\begin{paracol}{2}}
	{\end{paracol}}

\setlength{\columnseprule}{0.4pt}

\newenvironment{verses}[1]
	{\begin{flushleft} \begin{enumerate}[leftmargin=*] \setcounter{enumi}{#1}}
	{\end{enumerate} \end{flushleft}}

\newenvironment{versicles}{\par\leavevmode\parskip=0pt}{}

\begin{document}

\chapter*{Gratiarum Actio Post Missam}

\begin{latinenglishsection}

\latinenglish{
	\section*{Canticle of David (3. 57-88, 56)}
	
	\gresetinitiallines{1}
	\gregorioscore{trium_puerorum_intone}
	
	\gresetinitiallines{0}
	\gregorioscore{canticle_of_daniel_1}
	
	\begin{verses}{1}
	
	\item Benedícite Angeli \textbf{Dó}mino: * benedícite \textit{cæli} \textbf{Dó}mino.
	
	\item Benedícite aquæ omnes, quæ super cælos sunt, \textbf{Dó}mino: * benedícite omnes virtútes Dó\textit{mini} \textbf{Dó}mino.
	
	\item Benedícite sol et luna \textbf{Dó}mino: * benedícite stellæ \textit{cæli} \textbf{Dó}mino.
	
	\item Benedícite omnis imber et \textbf{Dó}mino: * benedícite omnes spíritus \textit{Dei} \textbf{Dó}mino.
	
	\item Benedícite ignis et æstus \textbf{Dó}mino: * benedícite frigus et \textit{æstus} \textbf{Dó}mino.
	
	\item Benedícite rores et pruína \textbf{Dó}mino: * benedícite gelu et \textit{fregus} \textbf{Dó}mino.
	
	\item Benedícite glácies et nives \textbf{Dó}mino: * benedícite noctes et \textit{dies} \textbf{Dó}mino.
	
	\item Benedícite lux et ténebræ \textbf{Dó}mino: * benedícite fúlgura et \textit{nubes} \textbf{Dó}mino.
	
	\item Benedícat terra \textbf{Dó}minum: * laudet et superexáltet e\textit{um in} \textbf{s\'{æ}}cula.
	
	\item Benedícite montes et colles \textbf{Dó}mino: * benedícite univérsa germinántia in \textit{terra} \textbf{Dó}mino.
	
	\item Benedícite fontes \textbf{Dó}mino: * benedícite mária et flú\textit{mina} \textbf{Dó}mino.
	
	\item Benedícite cete, et ómnia, quæ movéntur in aquis \textbf{Dó}mino: * benedícite omnes vólucres \textit{cæli} \textbf{Dó}mino.
	
	\item Benedícite omnes béstiæ et pécora \textbf{Dó}mino: * benedícite fílii hó\textit{minum} \textbf{Dó}mino.
	
	\item Benedícat Israel \textbf{Dó}minum: * laudet et superexáltet e\textit{um in} \textbf{sæ}cula.
	
	\item Benedícite sacerdótes \textbf{Dó}mino: * benedícite servi Dó\textit{mini} \textbf{Dó}mino.
	
	\item Benedícite spíritus, et ánimæ justórum \textbf{Dó}mino: * benedícite sancti et húmiles \textit{corde} \textbf{Dó}mino.
	
	\item Benedícite Ananía, Azaría, Mísael \textbf{Dó}mino: * laudáte et superexaltáte \textit{eum in} \textbf{s\'{æ}}cula.
	
	\item Benedícamus Patrem et Fílium cum Sancto \textbf{Spí}ritu: * laudémus et superexaltáte \textit{eum in} \textbf{s\'{æ}}cula.
	
	\item Benedíctus es Dómine in firmanénto \textbf{cæ}li: * et laudábilis, et gloriósus, et superexal\textit{tátus in} \textbf{s\'{æ}}cula.
	
	\end{verses}
	
	\section*{Psalm 150}
	
	\begin{verses}{0}
	
	\item Laudate Dóminum in sanctis \textbf{e}jus: * laudáte eum in firmaménto vir\textit{tútis} \textbf{e}jus.
	
	\item Laudáte eum in virtútibus \textbf{e}jus: * laudáte eum secúndum multitúdinem magnitú\textit{dinis} \textbf{e}jus.
	
	\item Laudáte eum in sono \textbf{tu}bæ: * laudáte eum in psaltéri\textit{o, et} \textbf{cí}thara.
	
	\item Laudáte eum in týmpano, et \textbf{cho}ro: * laudáte eum in chor\textit{dis, et} \textbf{ór}gano.
	
	\item Laudáte eum in cýmbalis benesonántibus: \dag\ laudáte eum in cýmbalis jubilati\textbf{ó}nis: * omnis spíritus \textit{laudet} \textbf{Dó}minum.
	
	\item Glória Patri, et \textbf{Fí}lio, * et Spirí\textit{tui} \textbf{Sanc}to.
	
	\item Sicut erat in princípio, et nunc, et \textbf{sem}per, * et in s\'{æ}cula sæcu\textit{lórum}. \textbf{A}men.
	
	\end{verses}
	
	\gregorioscore{trium_puerorum}
	
	\Vbar. Kyrie, eleison.
	\Rbar. Kyrie, eleison.
	\Vbar. Christe, eleison.
	\Rbar. Christe, eleison.
	\Vbar. Kyrie, eleison.
	\Rbar. Kyrie, eleison.
	\Vbar. Kyrie, eleison.
	
	\rubric{The first two words of the pater noster are intoned by the officiant, with the rest being said silently by all, until the officiant audibly chants the verse.}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) Pa(h)ter(h) no(h)ster(f.)... (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>V/</sp>. Et(h) ne(h) nos(h) in(h)dú(h)cas(h) in(h) ten(h)ta(h)ti(h)ó(h)nem(f.): (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>R/</sp>. Sed(h) lí(h)be(h)ra(h) nos(h) a(h) ma(h)lo(f.). (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>V/</sp>. Con(h)fi(h)te(h)án(h)tur(h) ti(h)bi(h), Dó(h)mi(h)ne(h), óm(h)ni(h)a(h) ó(h)pe(h)ra(h) tu(h)a(f.) . (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>R/</sp>. Et(h) San(h)cti(h) tu(h)i(h) be(h)ne(h)dí(h)cant(h) ti(h)bi(f.). (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>V/</sp>. Ex(h)sul(h)tá(h)bunt(h) San(h)cti(h) in(h) gló(h)ri(h)a(f.). (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>R/</sp>. Sed(h) nó(h)mi(h)ni(h) tu(h)o(h) da(h) gló(h)ri(h)am(f.). (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>V/</sp>. Dó(g)mi(h)ne(h), e(h)xáu(h)di(g) o(g)ra(g)ti(h)ó(h)nem(h) me(hg)am(g). (::)}
	
	\gresetinitiallines{0}
	\gabcsnippet{(c4) <sp>R/</sp>. Et(g) cla(h)mor(h) me(h)us(g) ad(g) te(g) vé(h)ni(h)at(h). (::)}
	
	Oremus. Deus, qui tribus púeris mitigásti flammas ígnium: \dag\ concéde propítius, ut nos fámulos tuos non exúrat flamma vitiórum.
	Actiones nostras, qu\'{æ}sumus Dómine, aspirándo pr\'{æ}veni, et adjuvándo proséquere: \dag\ ut cuncta nostra orátio et operátio a te semper incípiat, et per te coepta finiátur. Da nobis, qu\'{æ}sumus Dómine, vitiórum nostrórum flammas \textit{exstín}\textbf{gue}re, * qui beato Lauréntio tribuísti tormentórum suórum incéndia superá\textbf{re}. Per Christum, Dóminum nostrum.
	
	\Rbar. Amen.
	}{
	Let us sing the hymn...
	
	1. O all ye works of the Lord, bless the Lord: praise and exalt Him above all forever.
	
	2. O ye Angels of the Lord, bless the Lord: O ye heavens, bless the Lord.
	
	3. O all ye waters that are above the heavens, bless the Lord: O all ye powers of the Lord, bless the Lord.
	
	4. O ye sun and moon, bless the Lord: O ye stars of heaven, bless the Lord.
	
	5. O every shower and dew, bless ye the Lord: O all ye spirits of God, bless the Lord.
	
	6. O ye fire and heat, bles the Lord: O ye cold and heat, bless the Lord.
	
	7. O ye dews and hoar frosts, bless the Lord: O ye frost and cold, bless the Lord.
	
	8. O ye ice and snow, bless the Lord: O ye nights and days, bless the Lord.
	
	9. O ye light and darkness, bless the Lord: O ye lightnings and coulds, bless the Lord.
	
	10. O let the earth bless the Lord: let it praise and exalt Him above all forever.
	
	11. O ye mountains and hills, bless the Lord: O all ye things that spring up in thine earth, bless the Lord.
	
	12. O ye fountains, bless the Lord: O ye seas and rivers, bless the Lord.
	
	13. O ye whales and all that move in the waters, bless the Lord: O all ye fowls of the air, bless the Lord.
	
	14. O all ye beasts and cattle, bless the Lord: O ye sons of men, bless the Lord.
	
	15. O let Israel bless the Lord: let them praise and exalt Him above all forever.
	
	16. O ye priests of the Lord, bless the Lord: O ye servants of the Lord, bless the Lord.
	
	17. O ye spirits and souls of the just, bless the Lord: O ye holy and humble of heart, bless the Lord.
	
	18. O Ananais, Azarias, and Misael, bless ye the Lord: praise and exalt Him above all forever.
	
	19. Let us bless the Father and the Son, with the Holy Ghost: let us praise and exalt Him above all forever.
	
	20. Blessed art Thou, O Lord, in the firmament of heaven: and worthy of praise, and glorious, and exalted above all forever.
	
	1. Praise ye the Lord in His holy-places: praise ye Him in the firmament of His power.
	
	2. Praise ye Him for His mighty acts: praise ye Him according to the multitude of His greatness.
	
	3. Praise Him with sound of the trumpet. Praise HIm with psaltery and harp.
	
	4. Praise Him with timbrel and dance: praise Him with strings and organs.
	
	5. Praise Him on high-sounding cymbals: praise Him on cymbals of joy. Let every spirit praise the Lord.
	
	6. Glory be to the Father, and to the Son, and to the Holy Ghost: 
	
	7. As it was, is now, and ever shall be, worth without end. Amen.
	
	Let us sing the hymn of the three children, * which thes holy ones sang of old in the fiery furnace, giving praise to the Lord. (\textit{P. T.} Alleluia).
	
	\Vbar. Lord, have mercy.
	\Rbar. Lord, have mercy.
	\Vbar. Christ, have mercy.
	\Rbar. Christ, have mercy.
	\Vbar. Lord, have mercy.
	\Rbar. Lord, have mercy.
	\Vbar. Lord, have mercy. 
	
	Our Father...
	
	\Rbar. And lead us not into temptation.
	\Vbar. But deliver us from evil.
	
	\Rbar. O Lord, hear my prayer.
	\Vbar. And let my cry come unto Thee.
	
	Let us pray. O God, who didst ally the flames of fire for the three children, grant in Thy mercy that the flame of vice may not consume us, Thy servants. Prevent, we beseech Thee, O Lord, our actions by Thine inspirations, and further them by Thine assistance: that every word and work of ours may begin always from Thee and by Thee be likewise ended. Quench in us, we beseech Thee, O Lord, the flame of vice, even as Thou didst enable blessed Laurence to overcome his fire of sufferings. Through Christ our Lord.
	
	\Rbar. Amen.
	}

\end{latinenglishsection}

\end{document}